%!TEX encoding = UTF-8 Unicode
%!TEX root = ../galgas-book.tex

\chapitreTypePredefiniLabelIndex{bool}

An \galgas{@bool} object has a boolean value. The two keywords \galgas{true} and \galgas{false} belong to the \galgas{@bool} type, and denotes the true and false values. No constructor is defined for the \galgas{@bool} type.

\section{Readers}

\readerSansArgument{cString}
{bool}
{1.8.7}
{string}
{Returns a string representation of the receiver's value.}
{returns the \texttt{\textquotedbl true\textquotedbl} string if the receiver's value is true, and the \texttt{\textquotedbl false\textquotedbl} string otherwise.}







\readerSansArgument{ocString}
{bool}
{1.8.7}
{string}
{Returns a string representation of the receiver's value.}
{returns the \texttt{\textquotedbl YES\textquotedbl} string if the receiver's value is true, and the \texttt{\textquotedbl NO\textquotedbl} string otherwise.}




\readerSansArgument{sint}
{bool}
{1.9.4}
{sint}
{Returns the receiver's value in an \refTypePredefini{sint} (32-bit signed integer) object.}
{returns the \texttt{1S} \refTypePredefini{sint} value if the receiver's value is true, and the \texttt{0S} \refTypePredefini{sint} value otherwise.}




\readerSansArgument{sint64}
{bool}
{1.9.4}
{sint64}
{Returns the receiver's value in an \refTypePredefini{sint64} (64-bit signed integer) object.}
{returns the \texttt{1LS} \refTypePredefini{sint64} value if the receiver's value is true, and the \texttt{0LS} \refTypePredefini{sint64} value otherwise.}




\readerSansArgument{uint}
{bool}
{1.9.4}
{uint}
{Returns the receiver's value in an \refTypePredefini{uint} (32-bit unsigned integer) object.}
{returns the \texttt{1} \refTypePredefini{uint} value if the receiver's value is true, and the \texttt{0} \refTypePredefini{uint} value otherwise.}




\readerSansArgument{uint64}
{bool}
{1.9.4}
{uint64}
{Returns the receiver's value in an \refTypePredefini{uint64} (64-bit unsigned integer) object.}
{returns the \texttt{1L} \refTypePredefini{uint64} value if the receiver's value is true, and the \texttt{0S} \refTypePredefini{uint64} value otherwise.}




\section{Logical Operators}

The \galgas{@bool} type supports the three logical operators:\newline

\begin{tabular}{|c|c|}
\hline
\& & And \\
\hline
\textbar & Or \\
\hline
\textasciicircum   & Exclusive or \\
\hline
\end{tabular}

Theses operators require both arguments to be \galgas{@bool} objects and return an \galgas{@bool} object.\newline


The \galgas{@bool} type supports the logical unary operator:\newline

\begin{tabular}{|c|c|}
\hline
\galgas{not} & Complementation \\
\hline
\end{tabular}

This operator returns an \galgas{@bool} object.







\section{Comparison Operators}

The \galgas{@bool} type supports the six comparison operators:\newline

\begin{tabular}{|c|c|}
\hline
$==$ & Equality \\
\hline
$!=$ & Non Equality \\
\hline
$<$  & Strict Lower Than \\
\hline
$<=$  & Lower or Equal \\
\hline
$>$  & Strict Greater Than \\
\hline
$>=$  & Greater or Equal \\
\hline
\end{tabular}

Theses operators require both arguments to be \galgas{@bool} objects, and return a \galgas{@bool} object.


