%!TEX encoding = UTF-8 Unicode
%!TEX root = ../galgas-book.tex

\chapitreTypePredefiniLabelIndex{location}

An \galgas{@location} object value is a location in a source file. Objects of this type are useful for pointing out an error or a warning location.

\section{The \texttt{here} Keyword}

The \galgas{here} keyword indicates the current parsing location is the current source file. Assigning an \galgas{@location} object from the \galgas{here} keyword is a way for initializing an \galgas{@location} object:\newline

\texttt{@location currentLocation := here ;}

\section{Constructor}

\constructeurSansArgument{nowhere}
{location}
{2.1.2}
{location}
{Returns an \galgas{@location} that does not points out any location.}
{The returned object responds \galgas{true} to the \refReaderPage{location}{isNowhere}.}

\section{Readers}

\readerSansArgument{column}
{location}
{1.8.2}
{uint}
{Returns an \galgas{@uint} value containing the column of the receiver's value.}
{this reader raises a run-time error if the receiver's value responds \galgas{true} to the \refReaderPage{location}{isNowhere}.}


\readerSansArgument{isNowhere}
{location}
{2.1.2}
{bool}
{Returns an \galgas{@bool} value indicating whether the receiver'value points out a source location or does not.}
{this reader returns \galgas{true} if the receiver's value does not point out an actual location in a text source (i.e. it has been constructed using the nowhere constructor), and \galgas{false} if the receiver's value points out an actual location in a text source (i.e. it has been constructed using the \galgas{here} keyword.}


\readerSansArgument{line}
{location}
{1.8.2}
{uint}
{Returns an \galgas{@uint} value containing the line of the receiver's value.}
{this reader raises a run-time error if the receiver's value responds \galgas{true} to the \refReaderPage{location}{isNowhere}.}


\readerSansArgument{locationIndex}
{location}
{1.8.2}
{uint}
{Returns an \galgas{@uint} value containing the the offset from the the beginning of the source of the location defined by receiver's value.}
{this reader raises a run-time error if the receiver's value responds \galgas{true} to the \refReaderPage{location}{isNowhere}.}


\readerSansArgument{locationString}
{location}
{1.8.2}
{string}
{returns an \galgas{@string} object that contains a string representation of the location defined by receiver's value.}
{this reader raises a run-time error if the receiver's value responds \galgas{true} to the \refReaderPage{location}{isNowhere}.}
