%!TEX encoding = UTF-8 Unicode
%!TEX root = ../galgas-book.tex

\chapitreTypePredefiniLabelIndex{sint}

An \ggs+@sint+ object value is a 32-bit signed integer value. You can initialize an \ggs+@sint+ object from an 32-bit signed integer constant:\\

\begin{galgas}
@sint mySignedInteger = 123_456S
\end{galgas}


Note that a 32-bit signed integer constant is characterized by the 'S' suffix.


\section{Constructors}

\subsectionConstructor{min}{sint}

\begin{galgas}
constructor min -> @sint
\end{galgas}

Returns an \ggs+@sint+ object that the minimum value of the 32-bit signed range ($-2^{31}$).





\subsectionConstructor{max}{sint}

\begin{galgas}
constructor max -> @sint
\end{galgas}


Returns an \ggs+@sint+ object that the maximum value of the 32-bit signed range ($2^{31}-1$).



\section{Getters}

\subsectionGetter{bigint}{sint}

\begin{galgasbox}
getter bigint -> @bigint
\end{galgasbox}

Ce \emph{getter} permet de convertir un \ggs!@sint! en \ggs!@bigint!. Comme la plage des valeurs des \ggs!bigint! n'est limitée que par la mémoire disponible, il n'échoue jamais.

\begin{galgas}
  message [[-1234S bigint] string] + "\n" # -1234
\end{galgas}


\subsectionGetter{double}{sint}

\begin{galgas}
getter double -> @double
\end{galgas}

Returns the receiver's value converted in a \ggs+@double+ object. As a 32-bit integer value can always be converted in a \ggs+@double+ value, this getter never fails.





\subsectionGetter{sint64}{sint}

\begin{galgas}
getter sint64 -> @sint64
\end{galgas}

Returns the receiver's value in an \refTypePredefini{sint64} (64-bit signed integer) object. As a 32-bit signed value can always be converted in a 64-bit signed value, this getter never fails.

This getter is the only way to convert an \refTypePredefini{sint} object into an \refTypePredefini{sint64} object.





\subsectionGetter{string}{sint}

\begin{galgas}
getter string -> @string
\end{galgas}

Returns a decimal string representation of the receiver's value. For an hexadecimal string representation of the receiver's value, see \refGetterPage{uint}{hexString} and \refGetterPage{uint}{xString}.







\subsectionGetter{uint}{sint}

\begin{galgas}
getter uint -> @uint
\end{galgas}

Returns the receiver's value in an \refTypePredefini{uint} (32-bit unsigned integer) object. An error is raised is receiver's value is negative.

This getter is the only way to convert an \refTypePredefini{sint} object into an \refTypePredefini{uint} object.




\subsectionGetter{uint64}{sint}

\begin{galgas}
getter uint64 -> @uint64
\end{galgas}

Returns the receiver's value in an \refTypePredefini{uint64} (64-bit unsigned integer) object. An error is raised is receiver's value is negative.

This getter is the only way to convert an \refTypePredefini{sint} object into an \refTypePredefini{uint64} object.





\section{Incrementation and decrementation}

The \refTypePredefini{sint} supports incrementation and decrementation instructions.

\texttt{@sint n = ... ; n ++ ; \# Incrementation}

\texttt{@sint p = ... ; p -- ; \# Decrementation}\newline

The incrementation instruction raises an error if receiver's value is equal to $2^{31}-1$.\newline

The decrementation instruction raises an error if receiver's value is equal to $-2^{31}$.\newline

Note that incrementation and decrementation are not available within an expression.




\section{Arithmetic Operators}

The \ggs+@sint+ type supports the five arithmetic diadic operators:\newline

\begin{tabular}{|c|c|}
\hline
$+$ & Addition \\
\hline
$-$ & Substraction \\
\hline
$*$ & Multiplication \\
\hline
$/$ & Division \\
\hline
\ggs+mod+ & Modulo \\
\hline
\end{tabular}

Theses operators require both arguments to be \ggs+@sint+ objects.\newline

A run-time error is raised if the operation leads to a 32-bit signed overflow.

The \ggs+@sint+ type supports the following arithmetic unary operators:\newline

\begin{tabular}{|c|c|}
\hline
$+$ & No operation \\
\hline
$-$ & Negate \\
\hline
\end{tabular}

This operator returns the receiver's value (an \ggs+@sint+ object). A run-time error is raised if "-" operator is invoked on an object whose value is $-2^{31}$.






\section{Shift Operators}


The \ggs+@sint+ type supports right and left shift operators:\newline

\begin{tabular}{|c|c|}
\hline
$<<$ & Left shift \\
\hline
$>>$ & Right shift \\
\hline
\end{tabular}

Theses operators require the right argument to be \ggs+@sint+ object, and the left argument to be \ggs+@uint+ object.\newline

Note the right shift inserts a zero bit in the most significant bit location if the receiver's value is negative, and a one bit otherwise (it is a arithmetic right shift).\newline

The actual amount of the shift is the value of the right-hand operand masked by 31, i.e. the shift distance is always between 0 and 31.




\section{Logical Operators}

The \ggs+@sint+ type supports the three bit-wise logical operators:\newline

\begin{tabular}{|c|c|}
\hline
$\&$ & Bit-wise and \\
\hline
\textbar & Bit-wise or \\
\hline
\^\  & Bit-wise exclusive or \\
\hline
\end{tabular}

Theses operators require both arguments to be \ggs+@sint+ objects.\newline


The \ggs+@sint+ type supports the bit-wise logical unary operator:\newline

\begin{tabular}{|c|c|}
\hline
$\sim$ & Bit-wise complementation \\
\hline
\end{tabular}

This operator returns an \ggs+@sint+ object.







\section{Comparison Operators}

The \ggs+@sint+ type supports the six comparison operators:\newline

\begin{tabular}{|c|c|}
\hline
$=$ & Equality \\
\hline
$!=$ & Non Equality \\
\hline
$<$  & Strict Lower Than \\
\hline
$<=$  & Lower or Equal \\
\hline
$>$  & Strict Greater Than \\
\hline
$>=$  & Greater or Equal \\
\hline
\end{tabular}

Theses operators require both arguments to be \ggs+@sint+ objects, and return a \ggs+@bool+ object.


