%!TEX encoding = UTF-8 Unicode
%!TEX root = ../galgas-book.tex

\chapitreTypePredefiniLabelIndex{bool}

An \ggs+@bool+ object has a boolean value. The two keywords \ggs+true+ and \ggs+false+ belong to the \ggs+@bool+ type, and denotes the true and false values. No constructor is defined for the \ggs+@bool+ type.

\section{Getters}

\subsectionGetter{cString}{bool}

\begin{galgas}
getter cString -> @string
\end{galgas}

Returns a string representation of the receiver's value: the \ggs+"true"+ string if the receiver's value is true, and the \ggs+"false"+ string otherwise.







\subsectionGetter{ocString}{bool}

\begin{galgas}
getter ocString -> @string
\end{galgas}

Returns a string representation of the receiver's value: the \ggs+"YES"+ string if the receiver's value is true, and the \ggs+"NO"+ string otherwise.




\subsectionGetter{sint}{bool}

\begin{galgas}
getter sint -> @sint
\end{galgas}

Returns the receiver's value in an \refTypePredefini{sint} (32-bit signed integer) object: the \ggs+1S+ \refTypePredefini{sint} value if the receiver's value is true, and the \ggs+0S+ \refTypePredefini{sint} value otherwise.




\subsectionGetter{sint64}{bool}

\begin{galgas}
getter sint64 -> @sint64
\end{galgas}

Returns the receiver's value in an \refTypePredefini{sint64} (64-bit signed integer) object: the ±ggs+1LS+ \refTypePredefini{sint64} value if the receiver's value is true, and the \ggs+0LS+ \refTypePredefini{sint64} value otherwise.




\subsectionGetter{uint}{bool}

\begin{galgas}
getter uint -> @iuint
\end{galgas}

Returns the receiver's value in an \refTypePredefini{uint} (32-bit unsigned integer) object: the \ggs+1+ \refTypePredefini{uint} value if the receiver's value is true, and the \ggs+0+ \refTypePredefini{uint} value otherwise.




\subsectionGetter{uint64}{bool}

\begin{galgas}
getter uint64 -> @uint64
\end{galgas}

Returns the receiver's value in an \refTypePredefini{uint64} (64-bit unsigned integer) object: the \ggs+1L+ \refTypePredefini{uint64} value if the receiver's value is true, and the \ggs+0L+ \refTypePredefini{uint64} value otherwise.




\section{Logical Operators}

The \ggs+@bool+ type supports the three logical operators:\newline

\begin{tabular}{|c|c|}
\hline
\& & And \\
\hline
\textbar & Or \\
\hline
\textasciicircum   & Exclusive or \\
\hline
\end{tabular}

Theses operators require both arguments to be \ggs+@bool+ objects and return an \ggs+@bool+ object.\newline


The \ggs+@bool+ type supports the logical unary operator:\newline

\begin{tabular}{|c|c|}
\hline
\ggs+not+ & Complementation \\
\hline
\end{tabular}

This operator returns an \ggs+@bool+ object.







\section{Comparison Operators}

The \ggs+@bool+ type supports the six comparison operators:\newline

\begin{tabular}{|c|c|}
\hline
$==$ & Equality \\
\hline
$!=$ & Non Equality \\
\hline
$<$  & Strict Lower Than \\
\hline
$<=$  & Lower or Equal \\
\hline
$>$  & Strict Greater Than \\
\hline
$>=$  & Greater or Equal \\
\hline
\end{tabular}

Theses operators require both arguments to be \ggs+@bool+ objects, and return a \ggs+@bool+ object.


